### 2020-10-06 README.Rmd
Estructura general del proyecto

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Mesoscale Data Assimilation - A PhD Thesis"
```

# mesoda

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/paocorrales/mesoda/master?urlpath=rstudio)

This repository contains the code for my future thesis.

### How to cite

Please cite this compendium as:

> Corrales, (`r format(Sys.Date(), "%Y")`). _Compendium of R code for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`.

## Contents

The **analysis** directory contains:

  - [:file\_folder: paper](/analysis/chapters): R Markdown source document
    for manuscript. Includes code to reproduce the figures and tables
    generated by the analysis. 
  - [:file\_folder: data](/analysis/data): Links to data used in the analysis. The data is not included here.
  - [:file\_folder: figures](/analysis/figures): Plots and other
    illustrations
  - [:file\_folder:
    supplementary-materials](/analysis/supplementary-materials):
    Supplementary materials including notes and other documents
    prepared and collected during the analysis.

## How to run in your broswer or download and run locally

This research compendium has been developed using the statistical programming
language R. To work with the compendium, you will need
installed on your computer the [R software](https://cloud.r-project.org/)
itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

The simplest way to explore the text and code is to click on
[binder](https://mybinder.org/v2/gh/paocorrales/mesoda/master?urlpath=rstudio)
to open an instance of RStudio in your browser, which will have the
compendium files ready to work with. Binder uses rocker-project.org
Docker images to ensure a consistent and reproducible computational
environment. These Docker images can also be used locally.

You can download the compendium as a zip from from this URL:
[master.zip](/archive/master.zip). After unzipping:
- open the `.Rproj` file in RStudio
- run `devtools::install()` to ensure you have the packages this analysis depends on (also listed in the

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.