### 2021-09-07 poster2.Rmd
poster 2 - figura analisis

---
title: "Poster 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(mesoda)
library(metR)
library(tidyverse)
library(lubridate)
library(data.table)
library(here)
library(patchwork)
library(unglue)
library(knitr)
library(kableExtra)

map <- rnaturalearth::ne_states(country = c("argentina", "Brazil", "Chile", "Uruguay", "Paraguay", "Bolivia"), returnclass = "sf")

geom_mapa <- function(fill = NA) {
  list(geom_sf(data = map, fill = fill, color = "black", size = 0.1, inherit.aes = FALSE),
       # geom_sf(data = map_limitrofes, fill = fill, color = "black", size = 0.1, inherit.aes = FALSE),
       coord_sf(ylim = c(-42, -19), xlim = c(-76, -51)),
       scale_x_longitude(ticks = 5),
       scale_y_latitude(ticks = 5))
}

arg_topo <- metR::GetTopography(-75+360, -50+360, -20, -60, resolution = 1/10, 
                                file.dir = "~/mesoda/analysis/data/derived_data/")
arg_topo[, lon := metR::ConvertLongitude(lon)]

cordillera <- list(
  ggnewscale::new_scale_fill(),
  geom_contour_fill(data = arg_topo, aes(x = lon, y = lat,
                                         z = h),
                    breaks = seq(1500, 8000, by = 700)),
  scale_fill_gradient(low = "#E2E6E6", high = "#7E7E7E", guide = "none",
                      oob = scales::squish))

experiments <- c("E4", "E5", "E6", "E7")
colores_exp <- c(E2 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", E8 = "#CC3311")

satinfo <- fread(here("analysis", "data", "derived_data", "satinfo.txt")) %>% 
  setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
derived_data <- "/home/paola.corrales/datosmunin3/EXP/derived_data/"
```

## Analysis

```{r summary-fields}

files <- Sys.glob(paste0(derived_data, "pw/pw_2*"))

dates <- c("20181122000000", "20181122060000")

pw <- purrr::map(files, function(f) {
  details <- unglue::unglue(basename(f), "pw_{date}_{exp}.nc")[[1]]

  if (details$date %in% dates) {

    ReadNetCDF(f, vars = c(value = "pw", lon = "XLONG", lat = "XLAT")) %>% 
      .[, `:=`(exp = details$exp, 
               date = lubridate::ymd_hms(details$date))] %>%
      .[]
  }
}) %>% 
  rbindlist() %>% 
  .[, c("x", "y") := wrf_project(lon, lat)]

files <- c(Sys.glob(paste0(derived_data, "mucape/mcape_ana_E[2,5,6,8]_20181122000000.nc")),
           Sys.glob(paste0(derived_data, "mucape/mcape_ana_E[2,5,6,8]_20181122060000.nc")))

mucape <- purrr::map(files, function(f) {
  details <- unglue::unglue(basename(f), "mcape_ana_{exp}_{date}.nc")[[1]]

  if (details$date %in% dates) {

    ReadNetCDF(f, vars = c(value = "cape_2d", lon = "XLONG", lat = "XLAT")) %>% 
      .[, `:=`(exp = details$exp, 
               date = lubridate::ymd_hms(details$date))] %>%
      setnames(old = "mcape_mcin_lcl_lfc", new = "variable") %>% 
      .[]
  }
}) %>% 
  rbindlist()

fcsts <- expand_grid(fcsts = c("20181122000000", "20181122060000"),
                     exp = c("E2", "E5", "E6", "E8"))

ana <- lapply(seq_len(nrow(fcsts)), function(f) {

  tmp <- ReadNetCDF(paste0(derived_data, "/u_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(u = "uvmet")) %>% 
    .[, v := ReadNetCDF(paste0(derived_data, "/v_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(v = "uvmet"), out = "vector")] %>% 
    .[, q := ReadNetCDF(paste0(derived_data, "/q_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(q = "QVAPOR"), out = "vector")] %>%     .[, date :=  ymd_hms(fcsts[f, 1])] %>% 
    .[, exp := fcsts[f, 2]] %>%  
    .[]
}) %>% 
  rbindlist()

cor <- ana[bottom_top %in% c(1, 16)] %>% 
  dcast(date + exp + south_north + west_east ~ bottom_top, value.var = c("u", "v")) %>% 
  .[, cortante := Mag(u_16 - u_1, v_16 - v_1)] %>% 
  .[, ":="(u_1 = NULL, 
           u_16 = NULL,
           v_1 = NULL,
           v_16 = NULL)]

ana[bottom_top %between% c(1, 7), .(v_mean = mean(v)), 
    by = .(south_north, west_east, date, exp)] %>% 
  .[pw, on = .NATURAL] %>% 
  .[, date := factor(date)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = value, fill = stat(level_d)),
                    proj = norargentina_lambert) +
  colorspace::scale_fill_continuous_divergingx(super = ScaleDiscretised,
                                               palette = "PRGn", mid = 25,
                                               guide = guide_colorsteps(barwidth = 25,
                                                                        barheight = 0.5, 
                                                                        title.position = "left", 
                                                                        title.vjust = 1,
                                                                        frame.colour = "black")) +
  geom_contour2(aes(z = v_mean, size = stat(level)),
                proj = norargentina_lambert, color = "black", 
                breaks = c(-10, -5)) +
  scale_size_continuous(breaks = c(-10, -5), range = c(0.8, 0.4),
                        labels = c("-10", "-5")) +
  labs(fill = latex2exp::TeX("Precipitable\nwater ($Kgm^{-2}$)"), size = latex2exp::TeX("V ($ms^{-1})")) +
  cordillera +
  geom_mapa() +
  facet_grid(date ~ exp, 
             labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
                                         E6 = "SATWND", E8 = "RAD"),
                                 date = c("2018-11-22 00:00:00" = "00 UTC", "2018-11-22 06:00:00" = "06 UTC"))) +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3))

  # ana[bottom_top %between% c(1, 10), .(theta_mean = mean(theta)), 
  #     by = .(south_north, west_east, date, exp)] %>% 
  # .[pw, on = .NATURAL] %>% 
  # ggplot(aes(x, y)) +
  # geom_contour_fill(aes(z = theta_mean, fill = stat(level_d)),
  #                   proj = norargentina_lambert,
  #                   breaks = c(-Inf, seq(290, 314, 2), Inf), limits = c(-Inf, Inf)) +
  # scale_fill_distiller(super = ScaleDiscretised,
  #                      palette = "RdYlBu", direction = -1,
  #                      guide = guide_colorsteps(barwidth = 25,
  #                                               barheight = 0.5, 
  #                                               title.position = "left", 
  #                                               title.vjust = 1,
  #                                               frame.colour = "black")) +
  # labs(fill = "Potential\ntemperature (K)") +
  # cordillera +
  # geom_mapa() +
  # facet_grid(. ~ exp, 
  #            labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
  #                                        E6 = "SATWND", E8 = "RAD"))) +
  # theme_minimal() +
  # theme(legend.position = "bottom",
  #       panel.ontop = TRUE,
  #       panel.grid = element_line(linetype = 3),
  #       strip.text = element_blank(), 
  #       strip.background = element_blank()) +

mucape %>% 
  .[variable == "mcape"] %>% 
  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
  .[cor, on = .NATURAL] %>% 
  .[, date := factor(date)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = value, fill = stat(level_d)), 
                    proj = norargentina_lambert, limits = c(500, 6000)) +
  scale_fill_distiller(super = ScaleDiscretised,
                       palette = "YlOrRd", direction = 1,
                       guide = guide_colorsteps(barwidth = 25,
                                                barheight = 0.5, 
                                                title.position = "left", 
                                                title.vjust = 1,
                                                frame.colour = "black")) +
  geom_contour2(aes(z = cortante, size = stat(level)),
                proj = norargentina_lambert, color = "black", 
                breaks = c(15, 30)) +
  scale_size_continuous(breaks = c(15, 30), range = c(0.8, 0.4),
                        labels = c("15", "30")) +
  labs(fill = latex2exp::TeX("MCAPE ($JKg^{-1}$)"), size = latex2exp::TeX("Wind\nshear ($ms^{-1}$)")) +
  cordillera +
  geom_mapa() +
  facet_grid(date ~ exp, 
             labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
                                         E6 = "SATWND", E8 = "RAD"),
                                 date = c("2018-11-22 00:00:00" = "00 UTC", "2018-11-22 06:00:00" = "06 UTC"))) +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3)) 
```