### 2022-05-03 analysis/chapters/DA_goes16.Rmd
more organization

---
title: "Assimilating GOES-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(mesoda)
library(metR)
library(tidyverse)
library(lubridate)
library(data.table)
library(here)
library(patchwork)
library(unglue)
library(knitr)
library(kableExtra)
# source(here::here("temp_R", "goes_functions.R"))

map <- rnaturalearth::ne_states(country = c("argentina", "Brazil", "Chile", "Uruguay", "Paraguay", "Bolivia"), returnclass = "sf")

geom_mapa <- function() {
  list(geom_sf(data = map, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE),
       coord_sf(xlim = c(-76, -52), ylim = c(-41, -20)))
}

#dominio de interés
xlim <- c(2600, 3950)
ylim <- c(3700, 4750)
```

## Cloud mask

```{r}
  nc <- ncdf4::nc_open('~/datosmunin3/DATA/GOES16/ABI-L2-ACMF/OR_ABI-L2-ACMF-M3_G16_s20183240000342_e20183240011109_c20183240011270.nc')

cld_msk <- ReadNetCDF('~/datosmunin3/DATA/GOES16/ABI-L2-ACMF/OR_ABI-L2-ACMF-M3_G16_s20183240000342_e20183240011109_c20183240011270.nc', 
                      vars = 'BCM', 
                      subset = list(x = xlim,
                                    y = ylim)) %>% 
  .[, c("lon", "lat") := goes_projection(x, y, nc)]

cld_msk %>% 
  ggplot(aes(lon, lat)) +
  geom_point(aes(color = BCM == 0)) +
  scale_color_manual(values = c("darkgrey", "white"), guide = NULL) +
  geom_mapa() +
  labs(x = NULL, y = NULL,
       title = "Cloud Mask", subtitle = "Time: 2018-11-20 00 UTC")

```

## Channel 8

```{r}
f <- '~/datosmunin3/DATA/GOES16/ABI-L1b-RadF/2018/325/00/OR_ABI-L1b-RadF-M3C08_G16_s20183250000341_e20183250011108_c20183250011152.nc'

nc <- ncdf4::nc_open(f)

ch_8 <- ReadNetCDF(f, vars = c(rad = "Rad", 
                               qf = "DQF", 
                               sd = "std_dev_radiance_value_of_valid_pixels", 
                               pixel = "valid_pixel_count"), 
                   subset = list(x = xlim, y = ylim)) %>% 
  .[, rad := calculate_rad(rad, nc)] %>% 
  .[, tb := rad_to_tb(rad, nc)] %>% 
  .[, c("lon", "lat") := goes_projection(x, y, nc)]

ch_8 %>% 
  ggplot(aes(lon, lat)) +
  geom_point(aes(color = tb)) +
  scale_color_topes() +
  geom_mapa() +
  labs(x = NULL, y = NULL)

ch_8 %>% 
  ggplot(aes(lon, lat)) +
  geom_point(aes(color = tb)) +
  scale_color_topes() +
  geom_point(data = cld_msk[BCM == 1], color = "darkgrey", size = 0.9) +
  geom_mapa() +
  labs(x = NULL, y = NULL)

```

## Channel 10

```{r}

files <- Sys.glob("~/datosmunin3/DATA/GOES16/ABI-L1b-RadF/2018/324/*/OR_ABI-L1b-RadF-M3C10_G16*")

ch10 <- map(files, function(f){

  meta <- unglue(f, "/home/paola.corrales/datosmunin3/DATA/GOES16/ABI-L1b-RadF/2018/{day}/{hour}/OR_ABI-L1b-RadF-M3C10_G16_{file}.nc")

  d <- chron::month.day.year(324, c(month = 1, day = 1, year = 2018))

  nc <- ncdf4::nc_open(f)

  ReadNetCDF(f, vars = c(rad = "Rad", 
                         qf = "DQF", 
                         sd = "std_dev_radiance_value_of_valid_pixels"), 
             subset = list(x = xlim, y = ylim)) %>% 
    .[, time := make_datetime(d$year, d$month, d$day, meta[[1]][["hour"]])] %>% 
    .[, rad := calculate_rad(rad, nc)] %>% 
    .[, tb := rad_to_tb(rad, nc)] %>% 
    .[, c("lon", "lat") := goes_projection(x, y, nc)]

}) %>% 
  rbindlist()

```
## Standar deviation

```{r eval=FALSE, include=FALSE}

source("~/mesoda/analysis/scripts/ejemplo_focus.R")

ch10 %>%  
  # data.table::copy() %>% 
  .[, tb_std := with_matrix(tb ~ x | y, fun = rolling_2d_sd), 
    by = .(time)]

a %>%   
  .[tb_std >= 0.5] %>% 
  ggplot(aes(lon, lat)) +
  geom_point(aes(color = tb_std)) +
  scale_color_viridis_c() +
  geom_mapa()

fechas <- unique(ch10$time)

for(d in fechas) {

  ch10 %>%   
    .[tb_std >= 0.5 & time == d] %>% 
    ggplot(aes(lon, lat)) +
    geom_point(aes(color = tb_std), size = 0.5) +
    scale_color_viridis_c() +
    geom_mapa() +
    labs(title = "SDTB >= 0.5",
         subtitle = paste0("Time: ", as_datetime(d)))

  ggsave(paste0("std_",d,".png"))

}

# xlim <- c(2600, 3950)
# ylim <- c(3700, 4750)
# 
# tb_array <- matrix(a$tb, 
#                    nrow = 4750-3700+1, 
#                    ncol = 3950-2600+1,
#                    byrow = TRUE)
# 
# filled.contour(tb_array)
# 
# sd_array <- raster::focal(as(tb_array, "RasterLayer"),
#               w = matrix(1/9, ncol = 3, nrow = 3),
#               fun = sd,
#               na.rm = TRUE)
# 
# filled.contour(as(sd_array, "matrix"))

```

## Configuration

* 11/21 - 00 UTC
* Only 10 members (small test)
* Only assimilates ch 8 (upper level water vapor)
* Thinning 60 km

```{r}
diag <- fread("~/datosmunin3/EXP/prueba_abi/diag/diag_abi_g16_ges.ensmean.csv") %>% 
  .[, ":="(V5 = ConvertLongitude(V5),
           V2 = V2+6)]

colnames(diag) <- c("sensor", "channel", "freq", "lat", "lon", "peak", "press", "dhr", "tb_obs", "tbc", "tbcnob", "errinv", "qc", "usage", "emis", "tlapchn", "rzen", "razi", "rlnd", "rice", "rsnw", "rcld", "rcldp", paste0("pred", seq(12)))
```

### Observations minus background

```{r}
diag %>% 
  .[channel %in% c(8:10)] %>% 
  ggplot(aes(lon, lat)) +
  # geom_contour_fill(aes(z = tbc, fill = ..level..)) +
  geom_point(aes(color = tbc), size = 1) +
  scale_colour_divergent() +
  geom_mapa() +
  facet_wrap(~channel) +
  labs(x =  NULL, y = NULL, color = "O-B [TB]") +
  theme_minimal()

```
### Quality Control flags

* 2 = crtm issues
* 3 = gross check failed
* 5 = surface issues
* 7 = cloud contamination
* 50 = standar deviation to big
* 53 = issues with emissivity from surface

```{r}
diag %>% 
  .[channel %in% c(8:10)] %>% 
  ggplot(aes(lon, lat)) +
  geom_point(aes(color = factor(qc)), size = 0.9) +
  geom_mapa() +
  facet_wrap(~channel) +
  labs(x =  NULL, y = NULL, color = "QC") +
  theme_minimal()
```

## Assimilable observations

qc == 0, error !=0 and I shoud check the peakwt variable for outliers.

```{r}
diag %>% 
  .[channel %in% c(8:10) &
      errinv != 0 & qc == 0] %>% 
  ggplot(aes(lon, lat)) +
  geom_point(aes(color = errinv)) +
  geom_mapa() +
  facet_wrap(~channel) +
  labs(x =  NULL, y = NULL, color = "Error") +
  theme_minimal()
```

```{r}
diag %>% 
  .[channel %in% c(8:10) &
      errinv != 0 & qc == 0] %>% 
  ggplot(aes(lon, lat)) +
  # geom_contour_fill(aes(z = tbc, fill = ..level..)) +
  geom_point(aes(color = tbc), size = 1) +
  scale_colour_divergent() +
  geom_mapa() +
  facet_wrap(~channel) +
  labs(x =  NULL, y = NULL, color = "O-B [TB]") +
  theme_minimal()

```

usage flag = 4 on satinfo for the 3 channels. This means: assimilate without bias correction.

```{r}
diag %>% 
  .[channel %in% c(8:10)] %>% 
  melt(measure.vars = c("tbc", "tbcnob")) %>% 
  # .[variable == "tbcnob"] %>% 
  ggplot(aes(value)) + 
  geom_density(aes(color = variable)) +
  scale_color_discrete(labels = c("BC", "no BC")) +
  facet_wrap(~channel) +
  labs(x = "diff O-B", color = NULL) +
  theme_minimal()
```

## Quality control analysis

```{r}
files <- Sys.glob("~/datosmunin3/EXP/E10/ANA/*00/diagfiles/asim_abi_g16_*.ensmean")

diag <- map(files, function(f) {

  read_diag_rad(f, "E10") %>% 
    .[, channel := channel + 6]

}) %>% 
  rbindlist()

files <- Sys.glob("~/datosmunin3/EXP/EG2/ANA/*00/diagfiles/asim_abi_g16_*.ensmean")

diag_eg2 <- map(files, function(f) {

  read_diag_rad(f, "EG2") %>% 
    .[, channel := channel + 6]

}) %>% 
  rbindlist()

```

```{r}
rbind(diag, diag_eg2) %>% 
  .[channel %in% c(8:10), .N, by = .(qc, date, channel, exp)] %>% 
  ggplot(aes(date, N)) +
  geom_point(aes(color = factor(qc))) +
  geom_line(aes(color = factor(qc))) +
  # scale_color_manual(values = c("darkorange","grey40", "grey60", "grey80"),
  #                    labels = c("0" = "pass qc", 
  #                               "3" = "gross check",
  #                               "5" = "surface check",
  #                               "50" = "std dev")) +
  labs(x = NULL, y = NULL, color = NULL) +
  facet_grid(exp~channel,  scales = "free_y",
             labeller = labeller(channel = channels)) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

* **Gross check:** if tbc > errf reject obs. errf = min(3*errf, ermax_rad)

* **Standar Deviation:** toss data from channel 10 and surface channels if standar deviations for channel 10 is >= 0.5. The std dev for each pixel is calculated from the bt 3x3 pixel area. For detection of cloud border? what about areas where the surface have a impact?

* **Surface checks:** toss data depending on the channel and surface type.

```{r}

diag %>% 
  .[channel %in% c(8:10)] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  ggplot(aes(factor(channel))) +
  geom_bar(aes(fill = factor(qc))) +
  scale_fill_manual(values = c("darkorange","grey40", "grey60", "grey80"),
                    labels = c("0" = "pass qc", 
                               "3" = "gross check",
                               "5" = "surface check",
                               "50" = "std dev")) +
  facet_wrap(~date, ncol = 7) +
  labs(fill = NULL, x = "Channel", subtitle = "EG1") +
  theme_minimal() +
  theme(legend.position = "bottom")

```

```{r}

diag_eg2 %>% 
  .[channel %in% c(8:10)] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  ggplot(aes(factor(channel))) +
  geom_bar(aes(fill = factor(qc))) +
  scale_fill_manual(values = c("darkorange","grey40", "grey60", "grey80"),
                    labels = c("0" = "pass qc", 
                               "3" = "gross check",
                               "5" = "surface check",
                               "50" = "std dev")) +
  facet_wrap(~date, ncol = 7) +
  labs(fill = NULL, x = "Channel", subtitle = "EG2") +
  theme_minimal() +
  theme(legend.position = "bottom")

```

```{r}
diag %>% 
  .[channel %in% c(8:10)] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_point(aes(color = factor(qc)), size = 0.9) +
  geom_mapa() +
  scale_color_manual(values = c("darkorange","grey40", "grey60", "grey80"),
                     labels = c("0" = "pass qc", 
                                "3" = "gross check",
                                "5" = "surface check",
                                "50" = "std dev")) +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  facet_grid(channel~date) +
  labs(subtitle = "Assimilated obs", x =  NULL, y = NULL, color = "QC") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
diag_eg2 %>% 
  .[channel %in% c(8:10)] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_point(aes(color = factor(qc)), size = 0.9) +
  geom_mapa() +
  scale_color_manual(values = c("darkorange","grey40", "grey60", "grey80"),
                     labels = c("0" = "pass qc", 
                                "3" = "gross check",
                                "5" = "surface check",
                                "50" = "std dev")) +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  facet_grid(channel~date) +
  labs(subtitle = "Assimilated obs (EG2)", x =  NULL, y = NULL, color = "QC") +
  theme_minimal() +
  theme(legend.position = "bottom")

```

Looks like there are less observations rejected due to the standar deviation threshold on the last cycles. 00 UTC is 21 PM in Argentina, so it may be asociated to a diurnal cycle?

```{r}
diag %>% 
  .[channel %in% c(10)] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_point(aes(color = factor(qc)), size = 0.9) +
  geom_mapa() +
  scale_color_manual(values = c("grey50", "grey70", "grey80", "cyan4"),
                     labels = c("0" = "pass qc", 
                                "3" = "gross check",
                                "5" = "surface check",
                                "50" = "std dev")) +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  facet_wrap(~date, ncol = 4) +
  labs(subtitle = "Channel 10", x =  NULL, y = NULL, color = "QC") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
diag_eg2 %>% 
  .[channel %in% c(10)] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_point(aes(color = factor(qc)), size = 0.9) +
  geom_mapa() +
  scale_color_manual(values = c("grey50", "grey70", "grey80", "cyan4"),
                     labels = c("0" = "pass qc", 
                                "3" = "gross check",
                                "5" = "surface check",
                                "50" = "std dev")) +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  facet_wrap(~date, ncol = 4) +
  labs(subtitle = "Channel 10 (EG2)", x =  NULL, y = NULL, color = "QC") +
  theme_minimal() +
  theme(legend.position = "bottom")
```
Only a few observations are rejected at each cycle. These observations are located over The Andes (surface type = snow)

```{r}
diag[channel %in% c(8:10) & qc == 5] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_point(aes(color = factor(rsnw)), size = 0.9) +
  geom_mapa() +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  facet_grid(channel~date) +
  labs(subtitle = "QC = 5 (surface type)", 
       x =  NULL, y = NULL,
       color = "snow?") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
diag[channel %in% c(8:10) & qc == 3] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  .[, gross := min(3*errinv, 2)] %>% 
  ggplot(aes(tbc)) +
  geom_histogram() +
  # geom_point(aes(color = tbc), size = 0.9) +
  # scale_color_divergent() +
  # geom_mapa() +
  facet_grid(channel~date) +
  labs(x = "O-B", subtitle = "QC = 3 (Gross check)") +
  theme_minimal()
```

## Analysis minus guess

```{r}
files_ana <- Sys.glob("~/datosmunin3/EXP/E9/ANA/*00/analysis.mem001")[c(1:7)]

coord <- ReadNetCDF(files_ana[1], vars = c(lon = "XLONG", lat = "XLAT"), subset = list(bottom_top = 1))

levs <- c(200, 300, 400, 500, 600, 700, 800, 900, 1000)

ana <- map(files_ana, function(f) {

  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/E6/ANA/{date}/analysis.mem001")

  ReadNetCDF(f, vars = c(t_ana = "T", p = "P", "PB", q_ana = "QVAPOR",
                         lon = "XLONG", lat = "XLAT")) %>% 
    .[, ":="(t_ana = tk(t_ana, p + PB, T_BASE = 290) - 273.15,
             p = p + PB,
             PB = NULL)] %>% 
    .[, date := ymd_hms(meta[[1]][["date"]])] %>% 
    .[, .(lev = levs,
          t_ana = approx(p, t_ana, xout = levs*100)$y,
          q_ana = approx(p, q_ana, xout = levs*100)$y),
      by = .(date, south_north, west_east)] %>%
    coord[., on = .NATURAL] %>% 
    .[, c("x", "y") := wrf_project(lon, lat)] %>% 
    .[]

}) %>% 
  rbindlist()

files_gue <- Sys.glob("~/datosmunin3/EXP/EG2/GUESS/*00/wrfarw.mem001")

gue <- map(files_gue, function(f) {

  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/EG2/GUESS/{date}/wrfarw.mem001")

  ReadNetCDF(f, vars = c(t_gue = "T", p = "P", "PB", q_gue = "QVAPOR",
                         lon = "XLONG", lat = "XLAT")) %>% 
    .[, ":="(t_gue = tk(t_gue, p + PB, T_BASE = 290) - 273.15,
             p = p + PB,
             PB = NULL)] %>% 
    .[, date := ymd_hms(meta[[1]][["date"]])] %>% 
    .[, .(lev = levs,
          t_gue = approx(p, t_gue, xout = levs*100)$y,
          q_gue = approx(p, q_gue, xout = levs*100)$y),
      by = .(date, south_north, west_east)] %>%
    coord[., on = .NATURAL] %>% 
    .[, c("x", "y") := wrf_project(lon, lat)] %>% 
    .[]

}) %>% 
  rbindlist()

files_conv <- Sys.glob("~/datosmunin3/EXP/E2/ANA/*00/analysis.mem001")[1:7]

ana_conv <- map(files_conv, function(f) {

  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/E2/ANA/{date}/analysis.mem001")

  ReadNetCDF(f, vars = c(t_conv = "T", p = "P", "PB", q_conv = "QVAPOR",
                         lon = "XLONG", lat = "XLAT")) %>% 
    .[, ":="(t_conv = tk(t_conv, p + PB, T_BASE = 290) - 273.15,
             p = p + PB,
             PB = NULL)] %>% 
    .[, date := ymd_hms(meta[[1]][["date"]])] %>% 
    .[, .(lev = levs,
          t_conv = approx(p, t_conv, xout = levs*100)$y,
          q_conv = approx(p, q_conv, xout = levs*100)$y),
      by = .(date, south_north, west_east)] %>%
    coord[., on = .NATURAL] %>% 
    .[, c("x", "y") := wrf_project(lon, lat)] %>% 
    .[]

}) %>% 
  rbindlist()

files_conv <- Sys.glob("~/datosmunin3/EXP/*/ANA/20181121000000/analysis.ensmean")[c(1:5,7:8)]

ana_conv <- map(files_conv, function(f) {

  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/{run}/ANA/20181121000000/analysis.ensmean")

  ReadNetCDF(f, vars = c(t_conv = "T", p = "P", "PB", q_conv = "QVAPOR",
                         lon = "XLONG", lat = "XLAT")) %>% 
    .[, ":="(t_conv = tk(t_conv, p + PB, T_BASE = 290) - 273.15,
             p = p + PB,
             PB = NULL)] %>% 
    .[, date := meta[[1]][["run"]]] %>% 
    .[, .(lev = levs,
          t_conv = approx(p, t_conv, xout = levs*100)$y,
          q_conv = approx(p, q_conv, xout = levs*100)$y),
      by = .(date, south_north, west_east)] %>%
    coord[., on = .NATURAL] %>% 
    .[, c("x", "y") := wrf_project(lon, lat)] %>% 
    .[]

}) %>% 
  rbindlist()
```

```{r}
temp <- ana_conv %>% 
  dcast(x + y + lev ~ date, value.var = "q_conv") %>% 
  .[, ":="(aut = E5 - E2,
           satwnd = E6 - E5,
           rad_bad = E7 - E6,
           rad = E8 - E6,
           # abi_old = prueba_abi - E2,
           abi = EG2 - E6,
           abi_sat = EG2 - E5,
           abi_sat_aut = EG2 - E2)] %>% 
  melt(measure.vars = c("aut", "satwnd", "rad_bad", "rad", "abi", "abi_sat", "abi_sat_aut"))

temp2 <- ana_conv %>% 
  dcast(x + y + lev ~ date, value.var = "t_conv") %>% 
  .[, ":="(aut = E5 - E2,
           satwnd = E6 - E5,
           rad_bad = E7 - E6,
           rad = E8 - E6,
           # abi_old = prueba_abi - E2,
           abi = EG2 - E6,
           abi_sat = EG2 - E5,
           abi_sat_aut = EG2 - E2)] %>% 
  melt(measure.vars = c("aut", "satwnd", "rad_bad", "rad", "abi", "abi_sat", "abi_sat_aut"))

for (i in 1:length(levs)) {
  temp %>% 
    .[lev == levs[i]] %>% 
    ggplot(aes(x, y)) +
    geom_contour_fill(aes(z = value*1000, fill = ..level..), proj = norargentina_lambert) +
    scale_fill_divergent_discretised(trans = mesoda::symlog_trans(),
                                     guide = guide_colorsteps(barhight = 10)) +
    geom_mapa() +
    labs(x = NULL, y = NULL, fill = NULL,
         title = "Humidity",
         subtitle = as.character(levs[i])) +
    facet_wrap(~variable, ncol = 4) +
    theme_minimal()

   ggsave(paste0("impact_2018112100_q_", levs[i], ".png"), height = 4)

  temp2 %>% 
    .[lev == levs[i]] %>% 
    ggplot(aes(x, y)) +
    geom_contour_fill(aes(z = value, fill = ..level..), proj = norargentina_lambert) +
    scale_fill_divergent_discretised(trans = mesoda::symlog_trans(),
                                     guide = guide_colorsteps(barhight = 10)) +
    geom_mapa() +
    labs(x = NULL, y = NULL, fill = NULL,
         title = "Temperature",
         subtitle = as.character(levs[i])) +
    facet_wrap(~variable, ncol = 4) +
    theme_minimal()

   ggsave(paste0("impact_2018112100t_", levs[i], ".png"), height = 4)

}
```

```{r}
dates <- unique(ana$date)

for (i in 1:length(dates)) {

  ana[gue, on = .NATURAL] %>% 
    .[date == dates[i]] %>% 
    .[, date := format(date, "%m/%d %H UTC")] %>%
    .[, a_g := t_ana - t_gue] %>%
    .[lev %in% c(300, 500, 700)] %>% 
    ggplot(aes(x, y)) +
    geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
    scale_fill_divergent_discretised(trans = mesoda::symlog_trans()) +
    geom_mapa() +
    labs(x = NULL, y = NULL, fill = "A-B",
         title = "Temperature",
         subtitle = as.character(dates[i])) +
    facet_grid(date~lev) +
    theme_minimal()

  ggsave(paste0("EG2_t_ab_", format(dates[i], "%Y-%m-%d_%H"), ".png"), height = 3)

  ana[gue, on = .NATURAL] %>% 
    .[date == dates[i]] %>% 
    .[, date := format(date, "%m/%d %H UTC")] %>%
    .[, a_g := q_ana - q_gue] %>%
    .[lev %in% c(300, 500, 700)] %>% 
    ggplot(aes(x, y)) +
    geom_contour_fill(aes(z = a_g*1000, fill = ..level..), proj = norargentina_lambert) +
    scale_fill_divergent_discretised(trans = mesoda::symlog_trans()) +
    geom_mapa() +
    labs(x = NULL, y = NULL, fill = "A-B",
         title = "Humidity",
         subtitle = as.character(dates[i])) +
    facet_grid(date~lev) +
    theme_minimal()

  ggsave(paste0("EG2_q_ab_", format(dates[i], "%Y-%m-%d_%H"), ".png"), height = 3)

}
```

```{r}

for (i in 1:length(levs)) {

  ana[gue, on = .NATURAL] %>% 
    .[lev == levs[i]] %>% 
    .[, date := format(date, "%m/%d %H UTC")] %>%
    .[, a_g := t_ana - t_gue] %>%
    ggplot(aes(x, y)) +
    geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
    scale_fill_divergent_discretised() +
    geom_mapa() +
    labs(x = NULL, y = NULL, fill = "A-B",
         title = "Temperature",
         subtitle = paste(as.character(levs[i]), "hPa")) +
    facet_wrap(~date, ncol = 4) +
    theme_minimal()

  ggsave(paste0("EG2_t_ab_", levs[i], ".png"), height = 4)

  ana[gue, on = .NATURAL] %>% 
    .[lev == levs[i]] %>% 
    .[, date := format(date, "%m/%d %H UTC")] %>%
    .[, a_g := q_ana - q_gue] %>%
    ggplot(aes(x, y)) +
    geom_contour_fill(aes(z = a_g*1000, fill = ..level..), proj = norargentina_lambert) +
    scale_fill_divergent_discretised() +
    geom_mapa() +
    labs(x = NULL, y = NULL, fill = "A-B",
         title = "Humidity",
         subtitle = paste(as.character(levs[i]), "hPa")) +
    facet_wrap(~date, ncol = 4) +
    theme_minimal()

  ggsave(paste0("EG2_q_ab_", levs[i], ".png"), height = 4)

}

```

```{r}
for (i in 1:length(levs)) {

  ana[ana_conv, on = .NATURAL] %>% 
    .[lev == levs[i]] %>% 
    .[, date := format(date, "%m/%d %H UTC")] %>%
    .[, a_g := t_ana - t_conv] %>%
    ggplot(aes(x, y)) +
    geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
    scale_fill_divergent_discretised() +
    geom_mapa() +
    labs(x = NULL, y = NULL, fill = "SATWND-CONV",
         title = "Temperature",
         subtitle = paste(as.character(levs[i]), "hPa")) +
    facet_wrap(~date, ncol = 4) +
    theme_minimal()

  ggsave(paste0("E62_t_ac_", levs[i], ".png"), height = 4)

  ana[ana_conv, on = .NATURAL] %>% 
    .[lev == levs[i]] %>% 
    .[, date := format(date, "%m/%d %H UTC")] %>%
    .[, a_g := q_ana - q_conv] %>%
    ggplot(aes(x, y)) +
    geom_contour_fill(aes(z = a_g*1000, fill = ..level..), proj = norargentina_lambert) +
    scale_fill_divergent_discretised() +
    geom_mapa() +
    labs(x = NULL, y = NULL, fill = "SATWND-CONV",
         title = "Humidity",
         subtitle = paste(as.character(levs[i]), "hPa")) +
    facet_wrap(~date, ncol = 4) +
    theme_minimal()

  ggsave(paste0("E62_q_ac_", levs[i], ".png"), height = 4)

}
```
```{r}
files_ana <- Sys.glob("~/datosmunin3/EXP/prueba_abi/ANA/*_*/analysis.mem001")[c(1,3:5)]

ana <- map(files_ana, function(f) {

  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/prueba_abi/ANA/20181121000000_{run}/analysis.mem001")

  ReadNetCDF(f, vars = c(t_ana = "T", p = "P", "PB", q_ana = "QVAPOR",
                         lon = "XLONG", lat = "XLAT")) %>% 
    .[, ":="(t_ana = tk(t_ana, p + PB, T_BASE = 290) - 273.15,
             p = p + PB,
             PB = NULL)] %>% 
    .[, run := meta[[1]][["run"]]] %>% 
    .[, .(lev = levs,
          t_ana = approx(p, t_ana, xout = levs*100)$y,
          q_ana = approx(p, q_ana, xout = levs*100)$y),
      by = .(run, south_north, west_east)] %>%
    coord[., on = .NATURAL] %>% 
    .[, c("x", "y") := wrf_project(lon, lat)] %>% 
    .[]

}) %>% 
  rbindlist()

ana %>% 
  .[lev == levs[i]] %>% 
  # .[, date := format(date, "%m/%d %H UTC")] %>%
  # .[, a_g := q_ana - q_conv] %>%
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = q_ana*1000, fill = ..level..), proj = norargentina_lambert) +
  # scale_fill_divergent_discretised() +
  geom_mapa() +
  labs(x = NULL, y = NULL, fill = NULL,
       title = "Humidity",
       subtitle = paste(as.character(levs[i]), "hPa")) +
  facet_wrap(~run, ncol = 4) +
  theme_minimal()

ana[run != "all"] %>% 
  dcast(x + y + lev ~ run, value.var = "q_ana") %>% 
  .[ana[run == "all"], on = .NATURAL] %>% 
  .[, ":="(all_8 = t_ana - ch08,
           all_9 = t_ana - ch09,
           all_0 = t_ana - ch10)] %>% 
  melt(measure.vars = c("all_8", "all_9", "all_0")) %>% 
  .[lev == levs[4]] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = value, fill = ..level..), proj = norargentina_lambert) +
  scale_fill_divergent_discretised() +
  geom_mapa() +
  labs(x = NULL, y = NULL, fill = NULL,
       title = "Humidity",
       subtitle = paste(as.character(levs[i]), "hPa")) +
  facet_wrap(~variable, ncol = 4) +
  theme_minimal()

```

```{r}
files <- Sys.glob("~/datosmunin3/EXP/prueba_abi/ANA/*00/diagfiles/asim_conv*.ensmean")

read_diag_conv(files, "EG1") %>% 
  .[type %in% c(120)] %>% 
  unique(., by = c("stationID", "lat", "lon", "date")) %>% 
  .[] %>% 
  ggplot(aes(ConvertLongitude(lon), lat)) + 
  geom_point() +
  geom_mapa() +
  labs(x = NULL, y = NULL, 
       title = "Soundings",
       subtitle = "2018-11-21 00 UTC")

```

```{r eval=FALSE, include=FALSE}
ana <- ReadNetCDF("~/datosmunin3/EXP/prueba_abi/ANA/20181120180000/analysis.mem001", 
                  vars = c(T_ana = "QVAPOR", lon = "XLONG", lat = "XLAT")) %>% 
  .[, c("x", "y") := wrf_project(lon, lat)]

gue <- ReadNetCDF("~/datosmunin3/EXP/prueba_abi/GUESS/20181120180000/wrfarw.mem001",
                  vars = c(T_guess = "QVAPOR", lon = "XLONG", lat = "XLAT")) %>% 
  .[, c("x", "y") := wrf_project(lon, lat)]

ana[gue, on = .NATURAL] %>% 
  .[, a_g := T_ana - T_guess] %>%
  .[bottom_top %in% c(2:7)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
  # scale_fill_divergent_discretised() +
  geom_mapa() +
  facet_wrap(~bottom_top) +
  theme_minimal()

```

```{r eval=FALSE, include=FALSE}
ana[gue, on = .NATURAL] %>% 
  .[, a_g := T_ana - T_guess] %>%
  .[bottom_top %in% c(10:19)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
  scale_fill_divergent_discretised() +
  geom_mapa() +
  facet_wrap(~bottom_top) +
  theme_minimal()
```
```{r eval=FALSE, include=FALSE}
ana[gue, on = .NATURAL] %>% 
  .[, ":="(level = fcase(bottom_top <= 10, "PBL",
                         bottom_top %between% c(11, 15), "low",
                         bottom_top %between% c(16, 20), "middle",
                         bottom_top %between% c(21, 26), "upper",
                         bottom_top %between% c(27, 36), "the rest"))] %>% 
  .[, a_g := T_ana - T_guess] %>%
  .[level %in% c("low", "middle", "upper")] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
  scale_fill_divergent_discretised() +
  geom_mapa() +
  labs(x = NULL, y = NULL, fill = "A-B") +
  facet_wrap(~level) +
  theme_minimal()

ana[gue, on = .NATURAL] %>% 
  .[, ":="(level = fcase(bottom_top <= 10, "PBL",
                         bottom_top %between% c(11, 15), "low",
                         bottom_top %between% c(16, 20), "middle",
                         bottom_top %between% c(21, 26), "upper",
                         bottom_top %between% c(27, 36), "the rest"))] %>% 
  .[, a_g := T_ana - T_guess] %>%
  .[bottom_top %in% c(13, 18, 25)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
  scale_fill_divergent_discretised() +
  geom_mapa() +
  labs(x = NULL, y = NULL, fill = "A-B") +
  facet_wrap(~bottom_top) +
  theme_minimal()

```

```{r eval=FALSE, include=FALSE}
ana[gue, on = .NATURAL] %>% 
  .[, a_g := T_ana - T_guess] %>%
  .[bottom_top > 10 & lon %between% c(-55, -50)] %>% 
  .[, .(mean_lat = mean(a_g)), by = .(lat, bottom_top)] %>% 
  ggplot(aes(lat, bottom_top)) +
  geom_contour_fill(aes(z = mean_lat, fill = stat(level))) +
  scale_fill_divergent_discretised()
```