### 2022-05-03 analysis/chapters/goes.Rmd

---
title: "Deterministic hires forecasts"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(mesoda)
library(metR)
library(tidyverse)
library(lubridate)
library(data.table)
library(here)
library(patchwork)
library(unglue)
library(tagger)
library(knitr)
library(kableExtra)
#source(here::here("temp_R", "goes_functions.R"))

map <- rnaturalearth::ne_states(country = c("argentina", "Brazil", "Chile", "Uruguay", "Paraguay", "Bolivia"), returnclass = "sf")

geom_mapa <- function() {
  list(geom_sf(data = map, fill = NA, color = "black", size = 0.1, inherit.aes = FALSE),
       coord_sf(xlim = c(-76, -52), ylim = c(-41, -20)))
}

#dominio de interés
xlim <- c(2600, 3950)
ylim <- c(3700, 4750)

lon_band <- c(-66.5, -61.5)
lat_band <- c(-35.5, -29)

scale_date <- function(ini = 20181120180000, end = 20181123120000,
                       break_bin = "6 hour") {
  scale_x_datetime(breaks = seq.POSIXt(ymd_hms(ini), ymd_hms(end), by = break_bin),
                   labels = function(x) {
                     fifelse(hour(x) == 0, format(x, "%H\n%b %d"), format(x, "%H"))
                   }) 
}

derived_data <- "/home/paola.corrales/datosmunin3/EXP/derived_data/"

colores_exp <- c(E2 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", 
                 E9 = "#CC3311", E10 = "chartreuse4")

channels <- c("8" = "Ch 8 - upper level", 
              "9" = "Ch 9 - middle level", 
              "10" = "Ch 10 - lower level")

satinfo <- fread("~/mesoda/analysis/data/derived_data/satinfo.txt") %>% 
  setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
```

```{r}
files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_1h_fcst_det_201811220*_E*_d02.csv")

fss_fcst <- purrr::map(files, function(f) {
  meta <- unglue(f, "/home/paola.corrales/mesoda/analysis/data/derived_data/fss_1h_fcst_det_{init}_{exp}_d02.csv")

  fread(f) %>% 
    .[, date := as_datetime(date)] %>% 
    .[, run := meta[[1]][["init"]]]
}) %>% 
  rbindlist()

# 
# files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_ana_ens_E*[1-9]*.csv")
# 
# fss_ana <- purrr::map(files, function(f) {
#   
#   fread(f) %>% 
#     .[, date := as_datetime(date)] %>% 
#     .[, run := "ana"]
# }) %>% 
#   rbindlist()

fss_fcst %>% 
  .[w %in% c(1, 5, 11) & q %in% c(1, 5, 25)] %>% 
  # .[date %between% c(ymd_hms(20181122000000), ymd_hms(20181123120000))] %>% 
  ggplot(aes(date, fss)) +
  geom_hline(yintercept = 1, color = "darkgray") +
  geom_line(aes(color = exp, linetype = run)) +
  # geom_point(aes(color = exp)) +
  scale_color_manual(values = c(colores_exp, EG3 = "blue"), 
                     labels = c(E2 = "CONV", E5 = "ASWS", 
                                E6 = "SATWND", EG3 = "only ABI",
                                E9 = "RAD", E10 = "ABI")) +
  scale_date(20181122000000, 20181123000000, "12 hours") +
  coord_cartesian(xlim = c(ymd_hms(20181122000000), ymd_hms(20181123120000))) +
  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 
                                            "10" = "10 mm", "25" = "25 mm"),
                                      w = c("1" = "2 km", "5" = "10 km", "11" = "20 km", 
                                            "51" = "500 km", "101" = "1000 km"))) +
  tagger::tag_facets(position = list(x = 0.05, y = 0.90)) +
  labs(color = NULL, x = NULL, y = "FSS", linetype = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom")

fss_fcst %>% 
  .[w %in% c(1, 5, 11) & q %in% c(1, 5, 25)] %>% 
  # .[date %between% c(ymd_hms(20181122000000), ymd_hms(20181123120000))] %>% 
  ggplot(aes(date, fss)) +
  geom_hline(yintercept = 1, color = "darkgray") +
  geom_line(aes(color = exp, linetype = run)) +
  # geom_point(aes(color = exp)) +
  scale_color_manual(values = c(colores_exp, EG3 = "blue"), 
                     labels = c(E2 = "CONV", E5 = "ASWS", 
                                E6 = "SATWND", EG3 = "only ABI",
                                E9 = "RAD", E10 = "ABI")) +
  scale_date(20181122000000, 20181123000000, "12 hours") +
  coord_cartesian(xlim = c(ymd_hms(20181122000000), ymd_hms(20181123100000))) +
  facet_grid(w + q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 
                                            "10" = "10 mm", "25" = "25 mm"),
                                      w = c("1" = "2 km", "5" = "10 km", "11" = "20 km", 
                                            "51" = "500 km", "101" = "1000 km"))) +
  tagger::tag_facets(position = list(x = 0.05, y = 0.90)) +
  labs(color = NULL, x = NULL, y = "FSS", linetype = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom")

```