### 2022-05-03 analysis/chapters/poster2.Rmd

---
title: "Poster 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(mesoda)
library(metR)
library(tidyverse)
library(lubridate)
library(data.table)
library(here)
library(patchwork)
library(unglue)
library(knitr)
library(kableExtra)

map <- rnaturalearth::ne_states(country = c("argentina", "Brazil", "Chile", "Uruguay", "Paraguay", "Bolivia"), returnclass = "sf")

geom_mapa <- function(fill = NA) {
  list(geom_sf(data = map, fill = fill, color = "black", size = 0.1, inherit.aes = FALSE),
       # geom_sf(data = map_limitrofes, fill = fill, color = "black", size = 0.1, inherit.aes = FALSE),
       coord_sf(ylim = c(-42, -19), xlim = c(-76, -50)),
       scale_x_longitude(ticks = 5),
       scale_y_latitude(ticks = 5))
}

arg_topo <- metR::GetTopography(-75+360, -50+360, -20, -60, resolution = 1/10, 
                                file.dir = "~/mesoda/analysis/data/derived_data/")
arg_topo[, lon := metR::ConvertLongitude(lon)]

cordillera <- list(
  ggnewscale::new_scale_fill(),
  geom_contour_fill(data = arg_topo, aes(x = lon, y = lat,
                                         z = h),
                    breaks = seq(1500, 8000, by = 700)),
  scale_fill_gradient(low = "#E2E6E6", high = "#7E7E7E", guide = "none",
                      oob = scales::squish))

scale_date <- function(ini = 20181120180000, end = 20181123120000,
                       break_bin = "6 hour") {
  scale_x_datetime(breaks = seq.POSIXt(ymd_hms(ini), ymd_hms(end), by = break_bin),
                   labels = function(x) {
                     fifelse(hour(x) == 0, format(x, "%H UTC\n%b %d"), format(x, "%H"))
                   }) 
}

experiments <- c("E4", "E5", "E6", "E7")
colores_exp <- c(E2 = "#0077BB", E5 = "#88CCEE", E6 = "#EE7733", E8 = "#CC3311")

satinfo <- fread(here("analysis", "data", "derived_data", "satinfo.txt")) %>% 
  setnames(c("!sensor/instr/sat", "chan"), c("sensor", "channel"))
derived_data <- "/home/paola.corrales/datosmunin3/EXP/derived_data/"
```

## Analysis

```{r summary-fields}

files <- Sys.glob(paste0(derived_data, "pw/pw_2*"))

dates <- c("20181122000000", "20181122060000")

pw <- purrr::map(files, function(f) {
  details <- unglue::unglue(basename(f), "pw_{date}_{exp}.nc")[[1]]

  if (details$date %in% dates) {

    ReadNetCDF(f, vars = c(value = "pw", lon = "XLONG", lat = "XLAT")) %>% 
      .[, `:=`(exp = details$exp, 
               date = lubridate::ymd_hms(details$date))] %>%
      .[]
  }
}) %>% 
  rbindlist() %>% 
  .[, c("x", "y") := wrf_project(lon, lat)]

files <- c(Sys.glob(paste0(derived_data, "mucape/mcape_ana_E[2,5,6,8]_20181122000000.nc")),
           Sys.glob(paste0(derived_data, "mucape/mcape_ana_E[2,5,6,8]_20181122060000.nc")))

mucape <- purrr::map(files, function(f) {
  details <- unglue::unglue(basename(f), "mcape_ana_{exp}_{date}.nc")[[1]]

  if (details$date %in% dates) {

    ReadNetCDF(f, vars = c(value = "cape_2d", lon = "XLONG", lat = "XLAT")) %>% 
      .[, `:=`(exp = details$exp, 
               date = lubridate::ymd_hms(details$date))] %>%
      setnames(old = "mcape_mcin_lcl_lfc", new = "variable") %>% 
      .[]
  }
}) %>% 
  rbindlist()

fcsts <- expand_grid(fcsts = c("20181122000000", "20181122060000"),
                     exp = c("E2", "E5", "E6", "E8"))

ana <- lapply(seq_len(nrow(fcsts)), function(f) {

  tmp <- ReadNetCDF(paste0(derived_data, "/u_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(u = "uvmet")) %>% 
    .[, v := ReadNetCDF(paste0(derived_data, "/v_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(v = "uvmet"), out = "vector")] %>% 
    .[, q := ReadNetCDF(paste0(derived_data, "/q_", fcsts[f, 1], "_", fcsts[f, 2], ".nc"), vars = c(q = "QVAPOR"), out = "vector")] %>%     .[, date :=  ymd_hms(fcsts[f, 1])] %>% 
    .[, exp := fcsts[f, 2]] %>%  
    .[]
}) %>% 
  rbindlist()

cor <- ana[bottom_top %in% c(1, 16)] %>% 
  dcast(date + exp + south_north + west_east ~ bottom_top, value.var = c("u", "v")) %>% 
  .[, cortante := Mag(u_16 - u_1, v_16 - v_1)] %>% 
  .[, ":="(u_1 = NULL, 
           u_16 = NULL,
           v_1 = NULL,
           v_16 = NULL)]

ana[bottom_top %between% c(1, 7), .(v_mean = mean(v)), 
    by = .(south_north, west_east, date, exp)] %>% 
  .[pw, on = .NATURAL] %>% 
  .[, date := factor(date)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = value, fill = stat(level_d)),
                    proj = norargentina_lambert) +
  colorspace::scale_fill_continuous_divergingx(super = ScaleDiscretised,
                                               palette = "PRGn", mid = 25,
                                               guide = guide_colorsteps(barwidth = 25,
                                                                        barheight = 0.5, 
                                                                        title.position = "left", 
                                                                        title.vjust = 1,
                                                                        frame.colour = "black")) +
  geom_contour2(aes(z = v_mean, size = stat(level)),
                proj = norargentina_lambert, color = "black", 
                breaks = c(-10, -5)) +
  scale_size_continuous(breaks = c(-10, -5), range = c(0.8, 0.4),
                        labels = c("-10", "-5")) +
  labs(fill = latex2exp::TeX("Precipitable\nwater ($Kgm^{-2}$)"), size = latex2exp::TeX("V ($ms^{-1})")) +
  cordillera +
  geom_mapa() +
  facet_grid(date ~ exp, 
             labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
                                         E6 = "SATWND", E8 = "RAD"),
                                 date = c("2018-11-22 00:00:00" = "00 UTC", "2018-11-22 06:00:00" = "06 UTC"))) +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 10),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3))

ggsave("pw_ana.png", width = 30, height = 20, units = "cm")

mucape %>% 
  .[variable == "mcape"] %>% 
  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
  .[cor, on = .NATURAL] %>% 
  .[, date := factor(date)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = value, fill = stat(level_d)), 
                    proj = norargentina_lambert, limits = c(500, 6000)) +
  scale_fill_distiller(super = ScaleDiscretised,
                       palette = "YlOrRd", direction = 1,
                       labels = function(x) JumpBy(x, 2, fill = ""),
                       guide = guide_colorsteps(barwidth = 25,
                                                barheight = 0.5, 
                                                title.position = "left", 
                                                title.vjust = 1,
                                                frame.colour = "black")) +
  geom_contour2(aes(z = cortante, size = stat(level)),
                proj = norargentina_lambert, color = "black", 
                breaks = c(15, 30)) +
  scale_size_continuous(breaks = c(15, 30), range = c(0.8, 0.4),
                        labels = c("15", "30")) +
  labs(fill = latex2exp::TeX("MCAPE ($JKg^{-1}$)"), size = latex2exp::TeX("Wind\nshear ($ms^{-1}$)")) +
  cordillera +
  geom_mapa() +
  facet_grid(date ~ exp, 
             labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
                                         E6 = "SATWND", E8 = "RAD"),
                                 date = c("2018-11-22 00:00:00" = "00 UTC", "2018-11-22 06:00:00" = "06 UTC"))) +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 10),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3))

ggsave("mcape_ana.png", width = 30, height = 20, units = "cm")
```
## Pronóstico

```{r}
files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_fcst_ens_*_E[2,5,6,8]_ens.csv")

fss <- purrr::map(files, function(f) {

  meta <- unglue(basename(f), "fss_6h_fcst_ens_{ini_date}_{exp}_ens.csv")

  fread(f) %>% 
    .[, date := as_datetime(date)] %>% 
    .[, ini_date := ymd_h(meta[[1]][["ini_date"]])]
}) %>% 
  rbindlist()

fss %>% 
  .[w %in% c(1, 11) & q %in% c(5, 25) & exp %in% c("E2", "E5", "E6", "E8")] %>% 
  ggplot(aes(date, fss)) +
  geom_hline(yintercept = 1, color = "darkgray") +
  geom_line(aes(color = exp, linetype = factor(ini_date))) +
  geom_vline(xintercept = c(as_datetime("2018-11-22 14:00:00"), as_datetime("2018-11-23 06:00:00")), 
             size = 0.5,
             color = "darkgrey") +
  # geom_point(aes(color = exp)) +
  scale_color_manual(values = colores_exp, labels = c(E2 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E8 = "RAD")) +
  scale_linetype_discrete(labels = c("2018-11-22 00:00:00" = "00 UTC", "2018-11-22 06:00:00" = "06 UTC")) +
  scale_date(20181122000000, 20181123120000, "12 hours") +
  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 
                                            "10" = "10 mm", "25" = "25 mm"),
                                      w = c("1" = "10 km", "5" = "50 km", 
                                            "11" = "100 km", 
                                            "51" = "500 km", "101" = "1000 km"))) +
  labs(color = NULL, x = NULL, y = "FSS", linetype = NULL) +
  theme_minimal()

ggsave("fss2.png", width = 15, height = 10, unit = "cm")
```

```{r}

values <- c(-19.5, -12, -7.5, -3, 1.5, 6, 9, 12, 15, 16.5, 18, 19.5, 21, 
            22.5, 24, 25.5, 27, 28.5, 30, 31.5, 33, 34.5, 36, 37.5, 39, 40.5, 
            42, 43.5, 45, 46.5, 48, 49.5, 51, 52.5, 54, 55.5, 57, 58.5, 60, 
            61.5, 63, 64.5, 66, 69, 76.5)

colours <- c("#3D4170", "#3D4E7C", "#3B5B84", "#3A6695", "#3672A4", "#3188BD", 
             "#2F89B9", "#2896C7", "#24A4CF", "#2AB1D6", "#53F336", "#4BE231", 
             "#44D12F", "#3FC12B", "#3AAF25", "#32A11F", "#2C881E", "#227216", 
             "#EEEF39", "#DEE539", "#D7DA35", "#CED130", "#C4C52B", "#CBAE1F", 
             "#D59716", "#DB8116", "#EC0A2E", "#CB001A", "#C20013", "#B10009", 
             "#990000", "#C3005D", "#D600A2", "#E901E9", "#CC00CC", "#B100B5", 
             "#9800A0", "#FEFFFF", "#DFF5EC", "#C5F1E0", "#B7ECD8", "#A6ECCF", 
             "#95E3C5", "#85E0BD")

coord <- ReadNetCDF("~/datosmunin3/EXP/E2/FCST/2018112200/NPP/NPP_2018-11-22_00_FC14.nc", vars = c("XLAT", "XLONG"))

# files <- c(Sys.glob("~/datosmunin3/EXP/E[2,5,6,8]/FCST/201811220*/NPP/NPP_2018-11-22_00_FC14.nc"),
#            Sys.glob("~/datosmunin3/EXP/E[2,5,6,8]/FCST/201811220*/NPP/NPP_2018-11-22_06_FC08.nc"))
# 
# pp <- purrr::map_df(files, function(f) {
#   
#   d <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/{exp}/FCST/{ini_date}/NPP/NPP_2018-11-22_{other}_FC{val_date}.nc")
#   
#   ReadNetCDF(f, vars = "PP") %>% 
#     .[, .(PP = mean(PP)), by = .(lon, lat)] %>%
#     .[, ":="(fecha_ini = d[[1]][["ini_date"]],
#              hora_val = d[[1]][["val_date"]],
#              exp = d[[1]][["exp"]])]
#   
# }) %>% 
#   coord[., on = .NATURAL] %>% 
#   .[, ":="(lon = XLONG, lat = XLAT)] %>% 
#   .[, ":="(XLONG = NULL, XLAT = NULL)] %>% 
#   .[, c("x", "y") := wrf_project(lon, lat)]

files <- Sys.glob("~/datosmunin3/EXP/derived_data/ppacum/E[2,5,6,8]_fcst_201811220[0,6]_mean_acum_1h.rds")

pp <- map_df(files, function(f) {

  d <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/derived_data/ppacum/{exp}_fcst_{ini_date}_mean_acum_1h.rds")

  if (d[[1]][["exp"]] %in% c("E2", "E8")) {

    read_rds(f) %>% 
      .[, ":="(fecha_ini = d[[1]][["ini_date"]],
               lead_time = date)]

  } else { 
    read_rds(f) %>% 
      .[, ":="(fecha_ini = d[[1]][["ini_date"]],
               date = lead_time)]

  }

}) %>% 
  .[lead_time %in% c(as_datetime("20181122140000"), as_datetime("20181123060000"))]

pp %>%  
  # .[lead_time == as_datetime("20181123060000")] %>%
  .[, lead_time := factor(lead_time)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = pp_acum, fill = stat(level)),  breaks = c(seq(1, 45, 5), Inf),
                    proj = norargentina_lambert) +
  geom_contour2(aes(z = pp_acum), color = "black", size = 0.02, breaks = seq(1, 45, 5)) +
  scale_fill_distiller(palette = "YlGnBu",
                       direction = 1,, 
                       labels = function(x) JumpBy(x, 3, fill = ""),
                       super = ScaleDiscretised,
                       guide = guide_colorsteps(barwidth = 25,
                                                barheight = 0.4, 
                                                title.position = "left", 
                                                title.vjust = 1,
                                                frame.colour = "black")) +
  geom_mapa() +
  # coord_sf(ylim = c(-42, -19), xlim = c(-76, -51)) +
  # scale_x_longitude(ticks = 5, labels = NULL) +
  labs(x = NULL, y = NULL, fill = "COLMAX (dBZ)") +
  facet_grid(lead_time + fecha_ini ~ exp, labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
                                                                      E6 = "SATWND", E8 = "RAD"),
                                                              fecha_ini = c("2018112200" = "00 UTC", "2018112206" = "06 UTC"))) +
  theme_minimal(base_size = 9) +
  theme(legend.position = "bottom",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3))

obs_pp <- readRDS("/home/paola.corrales/mesoda/analysis/data/derived_data/IMERG_1h.rds") %>% 
  .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
  .[x %between% c(-300000, 800000) & y %between% c(-800000, 1090000) & 
      end_date %in% c(as_datetime("20181122140000"), as_datetime("20181123060000"))]

obs_pp %>% 
  .[, end_date := factor(end_date)] %>% 
  # .[end_date == as_datetime("20181122140000")] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = pp_acum, fill = stat(level)),  breaks = c(seq(1, 45, 5), Inf),
                    proj = norargentina_lambert) +
  geom_contour2(aes(z = pp_acum), color = "black", size = 0.02, breaks = seq(1, 45, 5)) +
  scale_fill_distiller(palette = "YlGnBu",
                       direction = 1,, 
                       labels = function(x) JumpBy(x, 3, fill = ""),
                       super = ScaleDiscretised,
                       guide = guide_colorsteps(barwidth = 25,
                                                barheight = 0.4, 
                                                title.position = "left", 
                                                title.vjust = 1,
                                                frame.colour = "black")) +
  geom_mapa() +
  facet_wrap(~end_date) +
  labs(x = NULL, y = NULL, fill = latex2exp::TeX("Acumulated\nprecipitation ($mmh^{-1}$)")) +
  theme_minimal(base_size = 9) +
  theme(legend.position = "bottom",
        axis.text.x = element_blank(),
        plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3)) 
```

```{r}
files <- Sys.glob("~/datosmunin3/EXP/derived_data/ppacum/E[2,5,6,8]_fcst_201811220[0,6]_prop_acum_1h.rds")

pp <- map_df(files, function(f) {

  d <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/derived_data/ppacum/{exp}_fcst_{ini_date}_prop_acum_1h.rds")

  if (d[[1]][["exp"]] %in% c("E2", "E8")) {

    read_rds(f) %>% 
      .[, ":="(fecha_ini = d[[1]][["ini_date"]],
               lead_time = date)]

  } else { 
    read_rds(f) %>% 
      .[, ":="(fecha_ini = d[[1]][["ini_date"]],
               exp = d[[1]][["exp"]],
               date = lead_time)]

  }

}) %>% 
  .[lead_time %in% c(as_datetime("20181122140000"), as_datetime("20181123060000"))]

obs_pp <- readRDS("/home/paola.corrales/mesoda/analysis/data/derived_data/IMERG_1h.rds") %>% 
  .[, c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
  .[x %between% c(-300000, 800000) & y %between% c(-800000, 1090000) & 
      end_date %in% c(as_datetime("20181122140000"), as_datetime("20181123060000"))]

pp %>%  
  # .[lead_time == as_datetime("20181123060000")] %>%
  .[umbral == 5] %>% 
  .[, lead_time := factor(lead_time)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = prop*100, fill = stat(level)),  breaks = seq(10, 100, 10),
                    proj = norargentina_lambert) +
  scale_fill_distiller(palette = "Spectral",
                       direction = -1,
                       labels = function(x) JumpBy(x, 1, fill = ""),
                       super = ScaleDiscretised,
                       guide = guide_colorsteps(barwidth = 25,
                                                barheight = 0.4,
                                                title.position = "left",
                                                title.vjust = 1,
                                                frame.colour = "black")) +
  geom_contour2(data = obs_pp %>% .[, lead_time := end_date], aes(z = pp_acum), color = "red", size = 0.2, 
                breaks = c(5),
                proj = norargentina_lambert) +
  geom_mapa() +
  # coord_sf(ylim = c(-42, -19), xlim = c(-76, -51)) +
  # scale_x_longitude(ticks = 5, labels = NULL) +
  labs(x = NULL, y = NULL, fill = latex2exp::TeX("Probability of 5 $mmh^{-1}$\n precipitation")) +
  facet_grid(lead_time + fecha_ini ~ exp, 
             labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", E6 = "SATWND", E8 = "RAD"),
                                 fecha_ini = c("2018112200" = "FCST: 00 UTC", "2018112206" = " FCST: 06 UTC"),
                                 lead_time = c("2018-11-22 14:00:00" = "11/22 14:00", "2018-11-23 06:00:00" = "11/23 06:00"))) +
  theme_minimal(base_size = 9) +
  theme(legend.position = "bottom",
        plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3))

ggsave("pp.png", width = 20, height = 20, units = "cm")
```